---
title: "New features in {gt} 0.6.0"
format:
  html:
    code-fold: true
    code-summary: "Show the code"
editor: visual
execute: 
  echo: true
  warning: false
  message: false
---

```{r}
library(tidyverse)
library(here)
library(gt)
```

### New functions in {gt} version 0.6.0

This is from a YouTube tutorial on \[Posit PBC channel: Rich Iannone \|\| New features in {gt} 0.6.0! \|\| RStudio\]l(https://www.youtube.com/watch?v=F5TV9uWCJps)

-   `sub_missing()` *nee* `fmt_missing`\
-   `sub_zero()`\
-   `sub_small_vals()`\
-   `sub_large_vals()`

### 1. Replace Missing Values

-   `fmt_missing()` is now `sub_missing()`
-   Default value for missing_text is ---\
-   sub_missing( *columns = 1:2, missing_text = emo::ji("clock")* )

```{r}
exibble %>% 
  select(-row, -group) %>% 
  gt() %>% 
  sub_missing(
    columns = 1:2,
    missing_text = "missing"
  ) %>% 
  sub_missing(
    columns = 4:7,
    missing_text = "nothing"
  )
  
```

### 2. sub_zero()

-   replace zeros\
-   default value is *nil*

Using default value for `sub_zer0()`

```{r}
tbl <- dplyr::tibble(num = c(10^(-1:2), 0, 0, 10^(4:6)))

tbl %>% 
  gt() %>% 
  fmt_number(columns = num) %>% 
  sub_zero()
```

#### Substitute text for Zero

```{r}
tbl <- dplyr::tibble(num = c(10^(-1:2), 0, 0, 10^(4:6)))

tbl %>% 
  gt() %>% 
  fmt_number(columns = num) %>% 
  sub_zero(zero_text = md("**_zero_**"))
```

### 3. sub_small_vals()

-   default value is 0.01; replaces small value with \<0.01\
-   i.e. default value =
    -   `sub_small_vals(small_pattern ="<{x}")`

#### Table with default `sub_small_vals()`

```{r}
tbl2 <- tibble(num = c(10^(-4:2), 0, NA)) 


tbl2 %>% 
  gt() %>% 
  fmt_number(columns = num) %>% 
  sub_small_vals() %>% 
  
   tab_header(
    title = md("sub_small_vals() default"),
    subtitle = md("`sub_small_vals()`")
  )


```

#### Table without modification

```{r}
tbl2 <- tibble(num = c(10^(-4:2), 0, NA)) 


tbl2 %>% 
  gt() %>% 
  fmt_number(columns = num) %>% 
  sub_small_vals(small_pattern = "") %>% 
  
   tab_header(
    title = md("sub_small_vals() default"),
    subtitle = md('`sub_small_vals(small_pattern = "")`')
  )



```

#### Table with modification

-   Change num to -num\
-   sub_small_vals(small_pattern = "-")

```{r}
tbl2 <- tibble(num = c(10^(-4:2), 0, NA)) 


tbl2 %>% 
  mutate(num = -num) %>% 
  gt() %>% 
  fmt_number(columns = num) %>% 
  sub_small_vals(small_pattern = "-") %>% 
  
   tab_header(
    title = md("sub_small_vals() with negative numbers"),
    subtitle = md('`sub_small_vals(small_pattern = "-")`')
  )



```

#### Using *sub_small_vals()* designating a threshold to 3 significant figures.

```{r}
tbl2 %>% 
  gt() %>% 
 # fmt_number(columns = num) %>%  
  
  # ----Keep three significant figures  ----------------------
  
  fmt_number(columns = num, n_sigfig = 3) %>%  
  
  sub_small_vals(
    threshold = 0.0005,
    small_pattern = "small"
  ) %>% 
  
  tab_header(
    title = md("**sub_small_vals() with threshold;<br /><br /> num to 3 Significant figures**"),
    subtitle = md('`sub_small_vals(threshold = 0.0005`,<br /><br />`small_pattern = "small")`')
  )

```

### 4. Using *sub_large_vals()*

```{r}

tbl3 <- dplyr::tibble(num = c(0, NA, 10^(8:14))) 

tbl3 %>% 
  gt() %>% 
  fmt_number(columns = num) %>% 
  sub_large_vals()  %>% 

 
  tab_header(
    title = md("**sub_large_vals()**"),
    subtitle = md("default sub_large_vals()")
  )
```

#### Negative numbers

```{r}

tbl3 <- dplyr::tibble(num = c(0, NA, 10^(8:14))) 

tbl3 %>% 
  mutate(num = -num) %>% 
  gt() %>% 
  fmt_number(columns = num) %>% 
  sub_large_vals(sign = "-")  %>% 

 
  tab_header(
    title = md("**sub_large_vals() with negative numbers**"),
    subtitle = md('default sub_large_vals(sign = "-")')
  )
```

#### sub_large_vals() with thresholds

```{r}


tbl3 <- dplyr::tibble(num = c(0, NA, 10^(8:14))) 

tbl3 %>% 
  gt() %>% 
  fmt_number(columns = num) %>% 
  sub_large_vals(
    threshold = 5E10,
    large_pattern = md("**humongous**")
  )  %>% 

 
  tab_header(
    title = md("**sub_large_vals() with thresholds**"))
#   subtitle = md("default sub_large_vals()")
  
```
